% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amssymb}
%
\begin{document}
%
\mainmatter              % start of the contributions
%
\title{Case Study of Liveness Verification in IronFleet}
%
\author{Lucas Pe\~{n}a and Manasvi Saxena}
%
\institute{University of Illinois at Urbana-Champaign, \\
\email{\{lpena7, msaxena2\}@illinios.edu}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
  As more complicated software systems are becoming prevalent in today's world,
  more sophisticated program verification techniques are needed to reason about
  these. In particular, verification of liveness properties, or asserting that a
  property must occur, is traditionally difficult. In this paper, we survey the
  IronFleet methodology~\cite{ironfleet}, a technique for building distributed
  systems that are provably correct, as it applies to verification of liveness
  properties. We consider the embedding of the Temporal Logic of Actions used in
  IronFleet, and discuss some limitations in their implementation.
\end{abstract}
%
\section{Introduction}
%Talk about what safety/liveness are at a high level (not going into verification).

\subsection{Safety Verification}
\subsection{Liveness Verification}
Liveness verification has the same theoretical complexity as safety verification
(cite). However, in practice, it is much more difficult to verify liveness
properties. In contrast to safety, verifying liveness requires proving the
absence of an infinite counterexample. Clearly, it is fruitless to do any kind
of brute force search for the absence of such a counterexample. Instead,
specific techniques for verifying liveness properties have been developed. These
include BDD-based model checking (cite), converting liveness properties to
safety properties (cite), and simulation-based liveness checking (cite).

The IronFleet methodology uses a proof strategy based on Lamport's Temporal
Logic of Actions (TLA) that chains smaller proofs together to assert the
existence of some final condition (see
Section~\ref{sec:liveness-ironfleet}). Unlike the related work on liveness
verification, IronFleet is the first to verify nontrivial liveness properties on
a large-scale distributed system.
%
\section{IronFleet}
\subsection{Dafny}
\subsection{End-to-end Verification}
\subsection{Refinement}
\subsection{Liveness Verification in IronFleet}\label{sec:liveness-ironfleet}
\subsubsection{Limitations}
%
\section{Temporal Logic of Actions}
The Temporal Logic of Actions (TLA) is an extension of Linear Temporal Logic
introduced to reason about concurrent systems. TLA introduces ``primed''
variables, which represents the value of a variable in the next state. For
example, the TLA \textit{action} $$x' = x + y$$ specifies that the value of $x$
in the next state is equal to the value of $x + y$ in the current state. An
action is satisfied by a pair of states $\langle s, t \rangle$, where $s$ is a
valuation on the unprimed variables and $t$ is a valuation on the primed
variables. The notion of primed variables allow us to specify a rich class of
formulas not expressible in normal LTL.

First, we can represent the action $Unchanged\;f \equiv f' = f$ for any 
function $f$  without primed variables, which states that $f$ is a
stuttering step. Using this, we can also represent the derived actions
$[\mathcal A]_f$ and $\langle \mathcal A \rangle_f$ for any action
$\mathcal A$. These respectively represent that an action either satisfies $\mathcal A$ or
stutters (w.r.t. $f$), and that an action that satisfies $\mathcal A$
necessarily changes $f$. These are defined as follows:
\begin{align*}
  [\mathcal A]_f \equiv \mathcal A \vee Unchanged\;f \hspace{1.5cm}
  & \langle \mathcal A \rangle_f \equiv \mathcal A \wedge \neg(Unchanged\;f)
\end{align*}
TLA also introduces the action $Enabled\;\mathcal A$ which specifies that for
any $s$, there is a $t$ such that $\langle s, t \rangle$ satisfies $\mathcal A$.

TLA \textit{formulas} add the temporal operator $\square$, like in LTL. In
addition, quantification over formulas is allowed. Derived formulas such as
$\lozenge$ are as in LTL. Other derived formulas include $F \leadsto G$
(\textit{leads to}), WF$_f(\mathcal A)$ (\textit{weak fairness}), and
SF$_f(\mathcal A)$ (\textit{strong fairness}). These are defined as follows:
\begin{align*}
  F \leadsto G &\equiv \square(F \Rightarrow \lozenge G) \\
  \text{WF}_f(\mathcal A) &\equiv \square\lozenge\langle\mathcal A\rangle_f \vee \square\lozenge\neg Enabled\;\langle\mathcal A\rangle_f \\
  \text{SF}_f(\mathcal A) &\equiv \square\lozenge\langle\mathcal A\rangle_f \vee \lozenge\square\neg Enabled\;\langle\mathcal A\rangle_f
\end{align*}
For use in verification, these derived formulas can express most liveness
properties a user may want to express. Of course, for verification, we need
proof rules allowing us to reason about these formulas, which we present below.

\vspace*{0.5cm}
TODO: rules here
\vspace*{0.5cm}

These rules allow us to reason about both strong fairness and weak fairness,
both as a hypothesis and a conclusion. The use of these rules is demonstrated
via a simple program that nondeterministically chooses one of two variables to
increment ad infinitum. One can define a naive implementation of this program,
as well as a more complex one using semaphores.

The TLA rules above can be used to prove weak fairness of the first
implementation, strong fairness of the second, and it can even be used to prove
that the semaphore implementation is a refinement of the naive implementation.
Notably, all four of the above rules must be used for these proofs.


\section{Related Work}
See~\cite{other-liveness}.
%
\section{Conclusion}

%
% ---- Bibliography ----
%
\bibliographystyle{ieeetr}

\bibliography{references}

\end{document}
